### 2022-06-18 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@47e9da202a84e537f493b7741affa594f76d6fa6 🚀

%
Norma\,Forero Rocio Munoz%
%
\,\textsuperscript{1,2}\quad %
Renata L.\,Muylaert%
%
\,\textsuperscript{3}\quad %
Stephanie N.\,Seifert%
%
\,\textsuperscript{4}\quad %
Gregory F.\,Albery%
%
\,\textsuperscript{5}\quad %
Colin J.\,Carlson%
%
\,\textsuperscript{6,7,8}\quad %
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{1,2,‡}

\textsuperscript{1}\,Université de
Montréal\quad \textsuperscript{2}\,Québec Centre for Biodiversity
Sciences\quad \textsuperscript{3}\,Molecular Epidemiology and Public
Health Laboratory, School of Veterinary Science, Massey University, New
Zealand\quad \textsuperscript{4}\,Paul G. Allen School for Global
Health, Washington State University, Pullman, WA, United
States\quad \textsuperscript{5}\,Department of Biology, Georgetown
University, Washington, DC, USA\quad \textsuperscript{6}\,Department of
Biology, Georgetown University, Washington,
DC,\quad \textsuperscript{7}\,Center for Global Health Science and
Security, Georgetown University Medical Center, Washington, DC,
USA\quad \textsuperscript{8}\,Department of Microbiology and Immunology,
Georgetown University Medical Center, Washington, DC, USA

@@

\vfill
Driven by the need to understand the ecological factors involved in the
emergence of betacoronavirus (the genus causing the SARS and MERS
disease in human) through bat hosts, we develop an approach to the
assesment of spillover risk based on the Geographic Mosaic Theory of
Coevolution. In doing so, we provide a global mapping of the spillover
risk posed by betacoronaviruses, reflecting the fact that this risk has
many ecological and evolutionary origins. Our framework reveals that
reservoir richness alone, although a component of viral hazard, is not a
sufficiently integrative predictor of risk. We offer alternative
insights based on viral sharing, host compositional uniqueness, and host
phylogenetic diversity. By comparing our aggregated measure of risk to a
proxy for human density, namely the proportion of each pixel that is
covered by urban or built land, we provide a synthetic risk map,
allowing the identification of hotspots where the bat-betacoronavirus
system may originate novel viruses in humans.

@@

Spillover risk is multidimensional and complicated. Within a pool of
competent hosts, it is driven by a multiplicity of factors (Plowright et
al. 2017). Although proxies for the local richness of hosts is commonly
analysed (see \emph{e.g.} Anthony et al. 2017 for coronaviruses), there
is an argument to be made that species who are not (known as) competent
hosts of a specific virus genus may not factor into this (Plowright et
al. 2015), calling for species-level information. This is especially
true as competence data increases predictive power when the taxonomic
scope of hosts of a viral family increases (Becker et al. 2020, Mull et
al. 2022). Similarly, host species who share viruses at different rates
should be weighted accordingly (Albery et al. 2020). In mammals, key
functional traits (for which phylogeny is a reasonable proxy) are
determinants of the spillover potential (Olival et al. 2017); these
include, notably, body mass, and affinity for urban environments (Albery
et al. 2022). Finally, especially when the pool of potential hosts spans
the entire globe, there may be local host pools that are highly unique;
not having been observed in other locations, these can act on the
overall risk either by providing novel contact opportunities, reflecting
unique host-environment combinations (Engering et al. 2013), or
facilitating rapid evolutionary changes in specialism of their pathogens
(Agosta et al. 2010). In the specific case of generalist pathogens
(which is the case many viruses in the \emph{Betacoronavirus} genus, see
\emph{e.g.} MacLean et al. 2021), there is conceptual and empirical
support to the idea that these community-level mechanisms are even more
important in driving the overall risk (Power and Mitchell 2004).

In this paper, we examine the biogeographic structure of
bats-betacoronaviruses associations, based on a curated dataset of all
confirmed bat hosts of betacoronaviruses. By drawing on concepts from
the Geographic Mosaic Theory of Coevolution (GMTC; Thompson 2005), we
turn these associations into a spatially explicit additive mapping of
zoonotic risk components, revealing the extreme heterogeneity of risk at
the global scale. Explicitely framing the notion of spillover risk based
on propositions from the GMTC (which is to say, based on a framework
linking interactions between species to change within species) is a
novel idea, that should be relatively general. Indeed, it only assumes
the action of well described evolutionary mechanisms. The benefit of
this approach is to provide the potential for a more dynamic and nuanced
understanding of risk: not only on ecological timescales, but also by
providing clues about which areas can change over micro-evolutionary
timescales. This provides a way to look at spatial structure by
accounting for more notions than species richness/similarity, but also a
way to identify spatial areas of higher risk.

We identify the Amazon and South-Eastern Asia as hotspots where the
phylogenetic distinctiveness of \emph{Betacoronavirus} is the highest
(Anthony et al. 2017); surprisingly, current data suggest that viral
sharing between hosts is high in the Amazon and low in South-Eastern
Asia, which has the potential to result in different evolutionary
dynamics between these two regions, hinting at different futures for
their viral communities. This work is important both as a description of
the bats-betacoronaviruses complex, but also because more broadly, bats
are known reservoirs for a variety of emerging viruses and other
pathogens (Calisher et al. 2006, Melaun et al. 2014), making balancing
the needs for bat conservation and disease prevention most likely very
difficult and a source of human-wildlife conflicts, especially in more
densely populated areas (Stone et al. 2015, Rego et al. 2015).

@@

We downloaded the data on bats hosts of \emph{Betacoronavirus} assembled
by Becker et al. (2022) from
\texttt{https://www.viralemergence.org/betacov} on Apr.~2022, and
filtered it to ``known'' hosts (established before the emergence of
SARS-CoV-2) and ``novel'' hosts (confirmed through sampling and
competence assays since the emergence of SARS-CoV-2). The original
database was assembled by a combination of data mining and literature
surveys, including automated alerts on the ``bats'' and ``coronavirus''
keywords to identify novel empirical evidence of bats-betacoronaviruses
associations; this yielded a total of 126 known hosts, 47 of which were
novel hosts.

@@

Nevertheless, locally diverse and virus-rich bat communities could
represent an increased risk of spillover under climate change through
the creation of novel interactions (Carlson et al. 2022), and therefore
the diversity of \emph{Betacoronavirus} strains should similarly be
ccounted for. In fig.~\ref{fig:richness} (bottom), we contrast the
evolutionary distinctiveness of bats and viruses -- this reveals a
slightly different portrait than bat richness alone. Chiropterans can be
classified, from a macro-evolutionary standpoint, as Yangochiroptera in
the ``New World,'' and Yinpterochiroptera elsewhere (Teeling et al.
2005, Springer 2013). Specifically, we would expect that the so-called
``New World'' group of bats, being more evolutionary distinct, would
also have evolutionary distinct viruses. Indeed fig.~\ref{fig:richness}
(bottom) reveals it to be the case, and this region harbors a distinct
bat-betacoronaviruses complex. By contrast, South-Eastern Asia has a lot
of non-evolutionary distinct bats, but evolutionary-distinct viruses.

@@

Our study focuses largely on the biogeography of hosts. Yet, we know
that viruses with high host plasticity, that is, the ability of a given
virus to adapt to various taxonomic orders and ecological groups
(Kreuder Johnson et al. 2015), are more likely to amplify viral
spillover, followed by secondary human-to-human transmission, and
geographical spread (Hazarie et al. 2021). High viral host plasticity is
an especially important trait for RNA viruses like betacoronaviruses
(Haddad et al. 2021). Indeed, our analysis of viral sequences reveals
that Latin America is a hotspot of viral distinctiveness, suggesting
that this part of the bats-betacoronaviruses complex may be undergoing
independent evolutionary dynamics (related species sharing viruses that
are different from the rest of the global pool). The other hotspot of
viral distinctiveness is S.E. Asia, in which richness is high but
sharing is low; this suggests a different type of evolutionary dynamics
(unrelated viruses coevolving with evolutionarily distinct hosts,
generating high diversity locally, see \emph{e.g.} Latinne et al. 2020).
Both of these areas should be priority areas for sampling, especially
since Becker et al. (2022) advance that they harbor undiscovered hosts
of beta-coronaviruses. This diversity of hosts, and the mechanisms by
which the exchange of viruses occurs between species, is largely
affected by the local environmental conditions and environmental change.

@@

\leavevmode\hypertarget{ref-Albery2022UrbMam}{}%
Albery, G. F. et al. 2022. Urban-adapted mammal species have more known
pathogens. in press.

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

%
Norma\,Forero Rocio Munoz%
%
\,\textsuperscript{1,2,‡}\quad %
Renata L.\,Muylaert%
%
\,\textsuperscript{3}\quad %
Stephanie N.\,Seifert%
%
\,\textsuperscript{4}\quad %
Gregory F.\,Albery%
%
\,\textsuperscript{5}\quad %
Daniel J.\,Becker%
%
\,\textsuperscript{6}\quad %
Colin J.\,Carlson%
%
\,\textsuperscript{7,8,9,‡}\quad %
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{1,2,‡}

\textsuperscript{1}\,Université de
Montréal\quad \textsuperscript{2}\,Québec Centre for Biodiversity
Sciences\quad \textsuperscript{3}\,Molecular Epidemiology and Public
Health Laboratory, School of Veterinary Science, Massey University, New
Zealand\quad \textsuperscript{4}\,Paul G. Allen School for Global
Health, Washington State University, Pullman, WA, United
States\quad \textsuperscript{5}\,Department of Biology, Georgetown
University, Washington, DC, USA\quad \textsuperscript{6}\,Department of
Biology, University of Oklahoma, Norman, OK,
USA\quad \textsuperscript{7}\,Department of Biology, Georgetown
University, Washington, DC,\quad \textsuperscript{8}\,Center for Global
Health Science and Security, Georgetown University Medical Center,
Washington, DC, USA\quad \textsuperscript{9}\,Department of Microbiology
and Immunology, Georgetown University Medical Center, Washington, DC,
USA

@@

\vfill
Driven by the need to understand the ecological factors involved in the
emergence of \emph{Betacoronavirus} (the genus causing the SARS, MERS,
and COVID-19 in human) through bat hosts, we develop an approach to the
assesment of spillover risk based on the Geographic Mosaic Theory of
Coevolution. In doing so, we provide a global mapping of the spillover
risk posed by betacoronaviruses, reflecting the fact that this risk is
best understood through the multi-faceted prism of ecological and
evolutionary mechanisms. Our framework reveals that reservoir richness
alone, although a component of viral hazard, is not a sufficiently
integrative predictor of risk. We offer alternative insights based on
viral sharing, host compositional uniqueness, and host phylogenetic
diversity and phylogeographic regions. By comparing our aggregated
measure of risk to a proxy for human density, namely the proportion of
each pixel that is covered by urban or built land, we provide a
synthetic risk map, allowing the identification of hotspots where the
bat-betacoronavirus system may originate novel viruses in humans.

@@

Spillover risk is multidimensional and complicated. Within a pool of
competent hosts, it is driven by a multiplicity of factors (Plowright et
al. 2017). Although proxies for the local richness of hosts are commonly
analysed (Anthony et al. 2017, Ruiz-Aravena et al. 2022), there is an
argument to be made that species who are not (known as) competent hosts
of a specific virus genus may not factor into this (Plowright et al.
2015), calling for species-level information. This is especially true as
competence data increases predictive power when the taxonomic scope of
hosts of a viral family increases (Becker et al. 2020, Mull et al.
2022). Similarly, host species who share viruses at different rates
should be weighted accordingly (Albery et al. 2020). In mammals, key
functional traits (for which phylogeny is a reasonable proxy) are
determinants of the spillover potential of their viruses (Olival et al.
2017); these include, notably, body mass, life history, diet, and use
for human-occupied spaces (Albery et al. 2022). Finally, especially when
the pool of potential hosts spans the entire globe, there may be local
host pools that are highly unique; not having been observed in other
locations, these can act on the overall risk either by providing novel
contact opportunities, reflecting unique host-environment combinations
(Engering et al. 2013), or facilitating rapid evolutionary changes in
specialism of their pathogens (Agosta et al. 2010). In the specific case
of generalist pathogens (which is the case for many viruses in the
\emph{Betacoronavirus} genus, see \emph{e.g.} MacLean et al. 2021),
there is conceptual and empirical support to the idea that these
community-level mechanisms are even more important in driving the
overall risk (Power and Mitchell 2004). In this paper, we examine the
biogeographic structure of bats-betacoronaviruses associations, based on
a curated dataset of all confirmed bat hosts of betacoronaviruses, by
adopting an analysis framework grounded in concepts from the Geographic
Mosaic Theory of Coevolution (GMTC; Thompson 2005). The GMTC posits that
coevolutionary processes between pairs (Thompson 1994) or complexes
(Janzen 1980) of species are structured in space by the rippling effects
of abiotic conditions onto evolutionary mechansism, resulting in
spatially fragmented evolutionary dynamics, coupled only by
dispersal-related processes (Gomulkiewicz et al. 2000). In turn, these
spatially fragmented processes can lead taxonomically homogeneous
systems (in our case, the bats-betacoronaviruses complex) to have
different structure and dynamics of large spatial extents (see
\emph{e.g.} Price 2002).

The GMTC predicts a spatial fragmentation of coevolutionary dynamics
under the joint action of three processes (see notably Gomulkiewicz et
al. 2007), which all have the potential to act on outbreak potential,
pathogen tramission, and disease virulence (Parratt et al. 2016, Turner
et al. 2021). First hot/cold spots of coevolution can appear when the
intensity of \emph{interaction} (in terms of reciprocal fitness
consequences) varies spatially, because of \emph{e.g.} partial range
overlap between organisms (Nuismer et al. 2003); hot/cold spots are
notoriously difficult to identify in nature (Laine 2005), especially
when the structure of range overlap is complex, and the reciprocal
fitness effets are relatively low. Both of these conditions are met in
the bats-coronaviruses complex. Second, GMTC supposes the existence of
selection mosaics, wherein the intensity of \emph{selection} (in terms
of selective consequences of fitness effects) varies across space; the
strength of reciprocal selection responds to the biotic complexity of
the community (locally diverse hosts and viruses are more biotically
complex; Thrall et al. 2007) \emph{and} to the local favorability of the
environment (Hochberg and Baalen 1998). In a system with near global
distribution and variations in relative richness, like the
bats-betacoronaviruses complex, the GMTC predicts the establishment of
many different coevolutionary trajectories. The final GMCT process is
trait remixing, under which changes in coevolutionary dynamics happen
due to the arrival (or departure) of functional traits, through changes
in community composition due to invasions, meta-community dynamics, and
disperal; adopting an host-centric perspective, high viral sharing would
expose the host immune system to multiple new viral traits, thereby
changing the selection landscape for all viruses.

We turn the processes on the GMTC into definitions of spillover risk
from viruses to hosts (focusing on the bats-betacoronavirus complex),
with a specific emphasis on the potential to create independant
coevolutionary dynamics (and therefore spatial fragmentation in the
risk) through heterogeneity. Our components of risk are (i) the
phylogenetic diversity of hosts, representing the diversification
potential of the system (hotspots, selection mosaics), under the
assumption that bat diversification is not primarily driven by viruses
(Gorbunova et al. 2020); (ii) the local uniqueness of the bat community,
representing the potential for viruses to be exposed to novel host
traits in unusual combinations (cold spots, trait remixing); finally,
(iii) viral sharing among hosts, representing the potential for hosts to
be exposed to complex and functionaly diverse communities (trait
remixing, biotic complexity). We turn these predicates into a spatially
explicit additive mapping of zoonotic risk components, revealing the
extreme heterogeneity of risk at the global scale. Explicitely framing
the notion of spillover risk based on propositions from the GMTC (which
is to say, based on a framework linking interactions between species to
change within species) is a novel idea, that should be relatively
general. Indeed, it only assumes the action of well described
evolutionary mechanisms. The benefit of this approach is to provide the
potential for a more dynamic and nuanced understanding of risk: not only
on ecological timescales, but also by providing clues about which areas
can change over micro-evolutionary timescales. This provides a way to
look at spatial structure by accounting for more notions than species
richness/similarity, but also a way to identify spatial areas of higher
risk.

@@

We downloaded the data on bats hosts of \emph{Betacoronavirus} assembled
by Becker et al. (2022) from
\texttt{https://www.viralemergence.org/betacov} on Apr.~2022, and
filtered it to ``known'' hosts (established before the emergence of
SARS-CoV-2) and ``novel'' hosts (confirmed through sampling and
competence assays since the initial data collection). The original
database was assembled by a combination of data mining and literature
surveys, including automated alerts on the ``bats'' and ``coronavirus''
keywords to identify novel empirical evidence of bats-betacoronaviruses
associations; this yielded a total of 126 known hosts, 47 of which were
novel hosts.

@@

Nevertheless, locally diverse and virus-rich bat communities could
represent an increased risk of spillover under climate change through
the creation of novel interactions (Carlson et al. 2022), and therefore
the diversity of \emph{Betacoronavirus} strains should similarly be
ccounted for. In fig.~\ref{fig:richness} (bottom), we contrast the
evolutionary distinctiveness of bats and viruses -- this reveals a
slightly different portrait than bat richness alone. Chiropterans can be
classified, from a macro-evolutionary standpoint, as Yangochiroptera and
Yinpterochiroptera elsewhere (Teeling et al. 2005, Springer 2013).
Specifically, we would expect that the so-called ``New World'' group of
bats, being more evolutionary distinct, would also have evolutionary
distinct viruses. Indeed fig.~\ref{fig:richness} (bottom) reveals it to
be the case, and this region harbors a distinct bat-betacoronaviruses
complex. This can be explained by the fact that Yangochiroptera,
although not limited to the western hemisphere, contain the highly
diverse adaptive radiation in the Phyllostomidae (Villalobos and Arita
2010), which is restricted to the western hemisphere. By contrast,
South-Eastern Asia has a lot of non-evolutionary distinct bats, who
nevertheless hosted evolutionary-distinct viruses.

@@

Our study focuses largely on the biogeography of hosts. Specifically, we
identify the Amazon and South-Eastern Asia as hotspots where the
phylogenetic distinctiveness of \emph{Betacoronavirus} is the highest
(Anthony et al. 2017); surprisingly, current data suggest that viral
sharing between hosts is high in the Amazon and low in South-Eastern
Asia, which has the potential to result in different evolutionary
dynamics between these two regions, hinting at different futures for
their viral communities. This work is important both as a description of
the bats-betacoronaviruses complex, but also because more broadly, bats
are known reservoirs for a variety of emerging viruses and other
pathogens (Calisher et al. 2006, Melaun et al. 2014), making balancing
the needs for bat conservation and disease prevention most likely very
difficult and a source of human-wildlife conflicts, especially in more
densely populated areas (Stone et al. 2015, Rego et al. 2015).

Yet, we know that viruses with high host plasticity, that is, the
ability of a given virus to adapt to various taxonomic orders and
ecological groups (Kreuder Johnson et al. 2015), are more likely to
amplify viral spillover, followed by secondary human-to-human
transmission, and geographical spread (Hazarie et al. 2021). High viral
host plasticity is an especially important trait for RNA viruses like
betacoronaviruses (Haddad et al. 2021). Indeed, our analysis of viral
sequences reveals that Latin America is a hotspot of viral
distinctiveness, suggesting that this part of the bats-betacoronaviruses
complex may be undergoing independent evolutionary dynamics (related
species sharing viruses that are different from the rest of the global
pool). The other hotspot of viral distinctiveness is S.E. Asia, in which
richness is high but sharing is low; this suggests a different type of
evolutionary dynamics (unrelated viruses coevolving with evolutionarily
distinct hosts, generating high diversity locally, see \emph{e.g.}
Latinne et al. 2020). Both of these areas should be priority areas for
sampling, especially since Becker et al. (2022) advance that they harbor
undiscovered hosts of beta-coronaviruses. This diversity of hosts, and
the mechanisms by which the exchange of viruses occurs between species,
is largely affected by the local environmental conditions and
environmental change.

@@

\leavevmode\hypertarget{ref-Albery2022UrbMam}{}%
Albery, G. F. et al. 2022. Urban-adapted mammal species have more known
pathogens. - Nature Ecology \& Evolution 6: 794--801.

@@

\leavevmode\hypertarget{ref-Gomulkiewicz2007DosDon}{}%
Gomulkiewicz, R. et al. 2007. Dos and don'ts of testing the geographic
mosaic theory of coevolution. - Heredity 98: 249--258.

@@

\leavevmode\hypertarget{ref-Hochberg1998AntCoe}{}%
Hochberg, M. E. and Baalen, M. 1998. Antagonistic coevolution over
productivity gradients. - The American Naturalist 152: 620--634.

@@

\leavevmode\hypertarget{ref-Janzen1980WheIt}{}%
Janzen, D. H. 1980. When is it Coevolution? - Evolution 34: 611--612.

@@

\leavevmode\hypertarget{ref-Laine2005SpaSca}{}%
Laine, A.-L. 2005. Spatial scale of local adaptation in a plant-pathogen
metapopulation. - Journal of Evolutionary Biology 18: 930--938.

@@

\leavevmode\hypertarget{ref-Nuismer2003CoeHos}{}%
Nuismer, S. L. et al. 2003. Coevolution between hosts and parasites with
partially overlapping geographic ranges. - Journal of Evolutionary
Biology 16: 1337--1345.

@@

\leavevmode\hypertarget{ref-Parratt2016InfDis}{}%
Parratt, S. R. et al. 2016. Infectious Disease Dynamics in Heterogeneous
Landscapes. - Annual Review of Ecology, Evolution, and Systematics 47:
283--306.

@@

\leavevmode\hypertarget{ref-Price2002MacThe}{}%
Price, P. W. 2002. Macroevolutionary Theory on Macroecological Patterns.
- Cambridge University Press.

@@

\leavevmode\hypertarget{ref-Ruiz-Aravena2022EcoEvo}{}%
Ruiz-Aravena, M. et al. 2022. Ecology, evolution and spillover of
coronaviruses from bats. - Nature Reviews Microbiology 20: 299--314.

@@

\leavevmode\hypertarget{ref-Thompson1994CoePro}{}%
Thompson, J. N. 1994. The Coevolutionary Process. - University of
Chicago Press.

@@

\leavevmode\hypertarget{ref-Thrall2007CoeSym}{}%
Thrall, P. H. et al. 2007. Coevolution of symbiotic mutualists and
parasites in a community context. - Trends in Ecology \& Evolution 22:
120--126.

\leavevmode\hypertarget{ref-Turner2021RolEnv}{}%
Turner, W. C. et al. 2021. The roles of environmental variation and
parasite survival in virulence--transmission relationships. - Royal
Society open science 8: 210088.

### 2022-06-18 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@47e9da202a84e537f493b7741affa594f76d6fa6 🚀

\medskip
\begin{flushleft}
{\small
%
Norma\,Forero Rocio Munoz%
%
\,\textsuperscript{1,2}, %
Renata L.\,Muylaert%
%
\,\textsuperscript{3}, %
Stephanie N.\,Seifert%
%
\,\textsuperscript{4}, %
Gregory F.\,Albery%
%
\,\textsuperscript{5}, %
Colin J.\,Carlson%
%
\,\textsuperscript{6,7,8}, %
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{1,2,‡}
\vskip 1em
\textsuperscript{1}\,Université de Montréal; \textsuperscript{2}\,Québec
Centre for Biodiversity Sciences; \textsuperscript{3}\,Molecular
Epidemiology and Public Health Laboratory, School of Veterinary Science,
Massey University, New Zealand; \textsuperscript{4}\,Paul G. Allen
School for Global Health, Washington State University, Pullman, WA,
United States; \textsuperscript{5}\,Department of Biology, Georgetown
University, Washington, DC, USA; \textsuperscript{6}\,Department of
Biology, Georgetown University, Washington,
DC,; \textsuperscript{7}\,Center for Global Health Science and Security,
Georgetown University Medical Center, Washington, DC,
USA; \textsuperscript{8}\,Department of Microbiology and Immunology,
Georgetown University Medical Center, Washington, DC, USA\\
\textsuperscript{‡}\,These authors contributed equally to the work\\
\vskip 1em
\textbf{Correspondance to:}\\
Timothée Poisot --- \texttt{timothee.poisot@umontreal.ca}\\
}
\end{flushleft}

@@

Driven by the need to understand the ecological factors involved in the
emergence of betacoronavirus (the genus causing the SARS and MERS
disease in human) through bat hosts, we develop an approach to the
assesment of spillover risk based on the Geographic Mosaic Theory of
Coevolution. In doing so, we provide a global mapping of the spillover
risk posed by betacoronaviruses, reflecting the fact that this risk has
many ecological and evolutionary origins. Our framework reveals that
reservoir richness alone, although a component of viral hazard, is not a
sufficiently integrative predictor of risk. We offer alternative
insights based on viral sharing, host compositional uniqueness, and host
phylogenetic diversity. By comparing our aggregated measure of risk to a
proxy for human density, namely the proportion of each pixel that is
covered by urban or built land, we provide a synthetic risk map,
allowing the identification of hotspots where the bat-betacoronavirus
system may originate novel viruses in humans.

>>>>

\medskip
\begin{flushleft}
{\small
%
Norma\,Forero Rocio Munoz%
%
\,\textsuperscript{1,2,‡}, %
Renata L.\,Muylaert%
%
\,\textsuperscript{3}, %
Stephanie N.\,Seifert%
%
\,\textsuperscript{4}, %
Gregory F.\,Albery%
%
\,\textsuperscript{5}, %
Daniel J.\,Becker%
%
\,\textsuperscript{6}, %
Colin J.\,Carlson%
%
\,\textsuperscript{7,8,9,‡}, %
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{1,2,‡}
\vskip 1em
\textsuperscript{1}\,Université de Montréal; \textsuperscript{2}\,Québec
Centre for Biodiversity Sciences; \textsuperscript{3}\,Molecular
Epidemiology and Public Health Laboratory, School of Veterinary Science,
Massey University, New Zealand; \textsuperscript{4}\,Paul G. Allen
School for Global Health, Washington State University, Pullman, WA,
United States; \textsuperscript{5}\,Department of Biology, Georgetown
University, Washington, DC, USA; \textsuperscript{6}\,Department of
Biology, University of Oklahoma, Norman, OK,
USA; \textsuperscript{7}\,Department of Biology, Georgetown University,
Washington, DC,; \textsuperscript{8}\,Center for Global Health Science
and Security, Georgetown University Medical Center, Washington, DC,
USA; \textsuperscript{9}\,Department of Microbiology and Immunology,
Georgetown University Medical Center, Washington, DC, USA\\
\textsuperscript{‡}\,These authors contributed equally to the work\\
\vskip 1em
\textbf{Correspondance to:}\\
Timothée Poisot --- \texttt{timothee.poisot@umontreal.ca}\\
}
\end{flushleft}

@@

Driven by the need to understand the ecological factors involved in the
emergence of \emph{Betacoronavirus} (the genus causing the SARS, MERS,
and COVID-19 in human) through bat hosts, we develop an approach to the
assesment of spillover risk based on the Geographic Mosaic Theory of
Coevolution. In doing so, we provide a global mapping of the spillover
risk posed by betacoronaviruses, reflecting the fact that this risk is
best understood through the multi-faceted prism of ecological and
evolutionary mechanisms. Our framework reveals that reservoir richness
alone, although a component of viral hazard, is not a sufficiently
integrative predictor of risk. We offer alternative insights based on
viral sharing, host compositional uniqueness, and host phylogenetic
diversity and phylogeographic regions. By comparing our aggregated
measure of risk to a proxy for human density, namely the proportion of
each pixel that is covered by urban or built land, we provide a
synthetic risk map, allowing the identification of hotspots where the
bat-betacoronavirus system may originate novel viruses in humans.
